### 2020-06-19 response/response-demography-3.tex

\section*{Deputy Editor}

\begin{enumerate}

\item The overall topic and research question is not clear in the
introduction. The main topic relates to changes in birth spacing over
time, which are influenced by sex selection/sex composition and maternal
education, and how this affects fertility and infant mortality. This is
a long list of interesting questions and analyses but this is confusing
for the reader. It would be useful to have a succinct statement about
what the study is about (what are key independent variables/concepts and
key outcomes) and place that in the introduction and before the data and
analyses section. This way, the reader can understand why the analyses
will proceed in a particular order and what is being tested. In
addition, the titles of the figures and tables should be reviewed to see
if they could better reflect the analyses undertaken. Overall, the
reader needs to be reminded at the beginning and in each section what
you aim to do.

[Response:] I have rewritten the Introduction to emphasize the main question I address
(``how did birth spacing respond to these changes, especially the introduction and spread 
of sex selection?'') and the two associated questions on the effects of the changes in 
birth spacing (``to what extent do changes in birth spacing bias our standard fertility 
estimates?'' and `` how do the changes in birth spacing and the increased use of abortions 
impact children's chances of survival?'').

@@

\item Please define sex selection early on.

[Response:] I now define ``sex selection'' as ``the selective abortion of female fetuses 
based on prenatal sex determination'' in the second paragraph of the paper.

### 2020-06-22 paper/sex_selection_spacing-ver4.tex
Minor edits of Introduction

India saw many positive changes over the last four decades:
the economy grew substantially,
schooling increased for both males and females,
and the total fertility rate fell to 2.2 
\citep{Bosworth2008,Dharmalingam2014,
International-Institute-for-Population-Sciences-IIPS2017}.

@@

The main question I examine is: how did birth spacing respond to these
changes, especially the introduction and spread of sex selection? 
The motivation is twofold.
First, prior research has failed to appreciate that each abortion increases the
interval between births by six to twelve months.%
\footnote{
The increase consists of three parts. 
First, after an abortion, the uterus needs at least two menstrual cycles to recover; 
otherwise, the likelihood of spontaneous abortion increases substantially \citep{zhou00b}. 
Second, the waiting time to conception is between one and six months. 
Finally, sex determination tests are reliable only from three months of gestation. 
}
Hence, the growing use of sex-selective abortions may substantially increase birth spacing, 
but we know little about how much. 
Second, more female education, higher household income, and low and declining female labor 
force participation are all also likely to impact birth spacing, and the combined changes 
may be much more substantial in India than what we have observed in other countries.

Understanding birth spacing is valuable because it contributes to our knowledge about 
fertility decisions and because of birth spacing's potential effects on outcomes such as 
fertility measures and mortality. 
I focus on two questions.

First, to what extent do changes in birth spacing bias our standard fertility estimates?
The tempo effect that arises from longer birth intervals means that the total fertility 
rate is a downward biased estimate of cohort fertility 
\citep{Hotz1997,Bongaarts1999,Ni-Bhrolchain2011}.
Hence, if birth intervals increased substantially over time, India's fertility may 
be higher than generally accepted.

@@

To examine how birth spacing has changed over time and across groups with the introduction 
of sex selection, I apply a competing risk hazard model to the birth histories of Hindu 
women, covering the period 1972 to 2016, using data from the four National Family and 
Health Surveys (NFHS).
The model directly incorporates the effects of sex-selective abortions on birth intervals
and the likelihood of a son, but it also works with no sex selection.
The empirical model allows me to predict fertility, taking into
account both the likelihood of parity progression and the likelihood of having a child of 
a specific sex.
I use the same data to examine how infant mortality risk has changed with the changing 
birth spacing and whether sex selection affects infant mortality.

% [Hypotheses / questions]
There are four main results.

>>>>

% Background
India saw many positive changes over the last four decades:
the economy grew substantially,
schooling increased for both males and females,
and the total fertility rate fell to 2.2 
\citep{Bosworth2008,Dharmalingam2014,
International-Institute-for-Population-Sciences-IIPS2017}.

@@

% Questions addressed and why valuable
The main question I examine is: how did birth spacing respond to these
changes, especially the introduction and spread of sex selection? 
The motivation is twofold.
First, prior research has failed to appreciate that each abortion increases the
interval between births by six to twelve months.%
\footnote{
The increase consists of three parts. 
First, after an abortion, the uterus needs at least two menstrual cycles to recover; 
otherwise, the likelihood of spontaneous abortion increases substantially \citep{zhou00b}. 
Second, the waiting time to conception is between one and six months. 
Finally, sex determination tests are reliable only from three months of gestation. 
}
Hence, the growing use of sex-selective abortions may substantially increase birth spacing, 
but we know little about how much. 
Second, more female education, higher household income, and low and declining female labor 
force participation are all also likely to impact birth spacing. 
Hence, the combined changes in birth spacing in India may be much more substantial than
 what we have observed in other countries.

Studying birth spacing is valuable because it contributes to our knowledge about 
fertility decisions and because of birth spacing's potential effects on outcomes such as 
fertility measures and mortality. 
To understand the effects of changes in birth spacing, I focus on two questions.

First, to what extent do changes in birth spacing bias our standard fertility estimates?
Tempo effects from longer birth intervals make the total fertility rate a downward biased 
estimate of cohort fertility \citep{Hotz1997,Bongaarts1999,Ni-Bhrolchain2011}.
Hence, if birth intervals increased substantially over time, India's fertility may 
be higher than generally accepted.

@@

% How the study progresses
To investigate how birth spacing has changed over time and across groups with the introduction 
of sex selection, I apply a competing risk hazard model to the birth histories of Hindu 
women, covering the period 1972 to 2016, using data from the four National Family and 
Health Surveys (NFHS).
The model directly incorporates the effects of sex-selective abortions on birth intervals
and the likelihood of a son, but it also works with no sex selection.
The empirical model allows me to predict fertility, taking into
account both the likelihood of parity progression and the likelihood of having a child of 
a specific sex.

I use the same data to examine how infant mortality risk has changed with the changing 
birth spacing and whether sex selection affects infant mortality.

% Main results
There are four main results.
